Emory Cardiovscular Biobank

### 2021-06-21 Manuscript/03-Simulation.Rmd
Add more content

$$
\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
$$
for stable poisson outcome. For the Gaussian outcome where dispersion are required, the dispersion parameter was set to 1. We follow the same data generating functions in Bai. In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active covariates, while the rest
covariates are inactive $f_j(x_j) = 0$ for $j = 4, \dots, p$. Another set of independent sample of size $n_{test}=1000$,($x_{new} ,y_{new}$), are created following the same data generating process, serving as the testing data. For mgcv and proposed models where choice of spline function are customizable, we choose to cubic regression spline with 10 basis. In COSSO and adaptive COSSO where the choice of spline are pre-determined, we set to have 10 bases. The corresponding number of parameter that we estimate with the model is 37, 82, 451 and 1801. We are run the simulation where the number of bases are increased to 25 deliberately. To evaluate the performance of the models, mean squared error, mean absolute error for gaussian model, AUC and misclassifcation rate are calculated for binomial model, and deviance are calculated for Poisson model, averaging across50 simulations. The results are tabulated in Table 1, 2, 3.

>>>>

$$
\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
$$
for stable poisson outcome. For the Gaussian outcome where dispersion are required, the dispersion parameter was set to 1. We follow the same data generating functions in Bai. In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active covariates, while the rest
covariates are inactive $f_j(x_j) = 0$ for $j = 4, \dots, p$. Another set of independent sample of size $n_{test}=1000$,($x_{new} ,y_{new}$), are created following the same data generating process, serving as the testing data. For mgcv and proposed models where choice of spline function are customizable, we choose to cubic regression spline with 10 basis. In COSSO and adaptive COSSO where the choice of spline are pre-determined, we set to have 10 bases. The corresponding number of parameter that we estimate with the model is 37, 82, 451 and 1801. We also run the simulation where the number of bases are increased to 25 deliberately. To evaluate the performance of the models, mean squared error, mean absolute error for gaussian model, AUC and misclassifcation rate are calculated for binomial model, and deviance are calculated for Poisson model, averaging across 50 simulations. The results are tabulated in Table 1, 2, 3.

```{r, fig.height=8}
targets::tar_load(binom_plot)
binom_plot
```

```{r, fig.height=8}
targets::tar_load(gaussian_plot)
gaussian_plot
```

```{r, fig.height=8}
targets::tar_load(poisson_plot)
poisson_plot
```

### 2021-06-21 Manuscript/SS_GAM.tex
Add more content

\[
\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
\] for stable poisson outcome. For the Gaussian outcome where dispersion
are required, the dispersion parameter was set to 1. We follow the same
data generating functions in Bai. In this data generating process, we
have \(x_1, x_2, x_3, x_4\) as the active covariates, while the rest
covariates are inactive \(f_j(x_j) = 0\) for \(j = 4, \dots, p\).
Another set of independent sample of size
\(n_{test}=1000\),(\(x_{new} ,y_{new}\)), are created following the same
data generating process, serving as the testing data. For mgcv and
proposed models where choice of spline function are customizable, we
choose to cubic regression spline with 10 basis. In COSSO and adaptive
COSSO where the choice of spline are pre-determined, we set to have 10
bases. The corresponding number of parameter that we estimate with the
model is 37, 82, 451 and 1801. We are run the simulation where the
number of bases are increased to 25 deliberately. To evaluate the
performance of the models, mean squared error, mean absolute error for
gaussian model, AUC and misclassifcation rate are calculated for
binomial model, and deviance are calculated for Poisson model, averaging
across50 simulations. The results are tabulated in Table 1, 2, 3.

>>>>

\[
\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
\] for stable poisson outcome. For the Gaussian outcome where dispersion
are required, the dispersion parameter was set to 1. We follow the same
data generating functions in Bai. In this data generating process, we
have \(x_1, x_2, x_3, x_4\) as the active covariates, while the rest
covariates are inactive \(f_j(x_j) = 0\) for \(j = 4, \dots, p\).
Another set of independent sample of size
\(n_{test}=1000\),(\(x_{new} ,y_{new}\)), are created following the same
data generating process, serving as the testing data. For mgcv and
proposed models where choice of spline function are customizable, we
choose to cubic regression spline with 10 basis. In COSSO and adaptive
COSSO where the choice of spline are pre-determined, we set to have 10
bases. The corresponding number of parameter that we estimate with the
model is 37, 82, 451 and 1801. We also run the simulation where the
number of bases are increased to 25 deliberately. To evaluate the
performance of the models, mean squared error, mean absolute error for
gaussian model, AUC and misclassifcation rate are calculated for
binomial model, and deviance are calculated for Poisson model, averaging
across 50 simulations. The results are tabulated in Table 1, 2, 3.

\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-8-1.pdf}

\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-9-1.pdf}

\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-10-1.pdf}

### 2021-06-21 Real_Data/Emory_Card_Biobank/Report.Rmd
Add more content

library(targets)
library(readr)
library(here)
library(gtsummary)
library(tidyverse)

train_dat <- targets::tar_read(RD_ECB_train_dat)
valid_dat <- targets::tar_read(RD_ECB_valid_dat)
```

@@

### Covariate Summary
```{r}
list(first_set = train_dat,
     second_set = valid_dat) %>% 
  map(~select(.x, # Data
              age, male, BlackRace,
           Strokehx:DM, 
           death3yr, timetodeath3yr) %>% 
        gtsummary::tbl_summary()
      ) %>% 
tbl_merge()
```

@@

@@

```{r}
library(BHAM)
library(BhGLM)

mz_names <- train_dat %>% select(starts_with("mz")) %>% names

test_mz_names <- valid_dat  %>% select(starts_with("mz")) %>% names

sm_df <- data.frame(
  Var = mz_names,
 Func = "s",
 Args ="bs='cr', k=5"
)

sm_obj <- construct_smooth_data(sm_df, train_dat)

dsn_mat <- sm_obj$data

# The validation data doesn't overlap with training data
# test_sm_dat <- make_predict_dat(sm_obj$Smooth, dat = valid_dat)

mdl_defaul <- bmlasso(dsn_mat, train_dat$death3yr, family = "binomial")

mdl_dflt <- bmlasso_spline(dsn_mat, train_dat$death3yr, family = "binomial",
                           group = make_group(names(dsn_mat)))

measure.bh(mdl_defaul)

@@

## Validation
```{r eval = F}

measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=valid_dat$death3yr)

>>>>

library(targets)
library(readr)
library(here)
library(gtsummary)
library(tidyverse)
library(BhGLM)
library(BHAM)

train_dat <- targets::tar_read(RD_ECB_train_dat)
valid_dat <- targets::tar_read(RD_ECB_valid_dat)

gam_scn_res <- targets::tar_read(train_gam_screen)
```

@@

### Covariate Summary
```{r eval = F}
list(first_set = train_dat,
     second_set = valid_dat) %>% 
  map(~select(.x, # Data
              age, male, BlackRace,
           Strokehx:DM, 
           death3yr, timetodeath3yr) %>% 
        gtsummary::tbl_summary()
      ) %>% 
tbl_merge()
```

@@

## Screening
### Linear Effect Screening
```{r eval = F}
# feature_name <- 
linear_screening_res <- train_dat %>% select(starts_with("mz")) %>% 
  names() %>% 
map_dfr(#feature_name[1:2],
        .f = function(name, .dat){
  # name <- feature_name[1]
  y <- .dat %>% pull(death3yr)
  x <- .dat %>% pull({{name}})
  glm(y ~ x, family = binomial()) %>% 
    tidy() %>% 
    filter(term == "x") %>% 
    select(p.value) %>% 
    mutate(var = name,
           p.value)

},
.dat = train_dat)

lnr_scn_res <- data.frame(
linear_screening_res,
p.adj = p.adjust(linear_screening_res$p.value, "fdr")
) %>% arrange(p.adj) %>% filter(p.adj<0.2) #%>% 
  # slice_head(n = 100)
```

### GAM Screening
```{r eval = F}

library(mgcv)
library(broom)

# feature_name <- 
gam_screening_res <- train_dat %>% select(starts_with("mz")) %>% 
  names() %>% 
map_dfr(#feature_name[1:2],
        .f = function(name, .dat){
  # name <- feature_name[1]
  y <- .dat %>% pull(death3yr)
  x <- .dat %>% pull({{name}})
  mgcv::gam(y ~ s(x, bs = "cr", k = 5), family = binomial(),
            data = data.frame(x, y)) %>% 
    tidy() %>% 
    filter(term == "s(x)") %>% 
    select(p.value) %>% 
    mutate(var = name,
           p.value)

},
.dat = train_dat)

gam_scn_res <- data.frame(
gam_screening_res,
p.adj = p.adjust(gam_screening_res$p.value, "fdr")
) %>% arrange(p.adj) %>% #filter(p.adj<0.2)
  head(200)
```

@@

```{r data_split, eval = F}
library(tidymodels)

set.seed(1)

data_split <- rsample::initial_split(train_dat, prop = 3/4)

train_data <- training(data_split)
test_data <- testing(data_split)
```

## Baseline: Linear Lasso Model

```{r eval = F}
library(glmnet)

train_data <- train_dat %>% select(gam_scn_res %>% pull(var), death3yr)

lasso_mdl_cv <- cv.glmnet(x = train_data %>% select(starts_with("mz")) %>% data.matrix,
                    y = train_data %>% pull(death3yr), alpha = 1,
                    family = "binomial", type.measure = "auc")

lasso_mdl <- glmNet(x = train_data %>% select(starts_with("mz")) %>% data.matrix,
                    y = train_data %>% pull(death3yr), alpha = 1,
                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)

lasso_cv_res <- cv.bh(lasso_mdl, ncv = 10)$measures

# class(lasso_mdl) <- c("glmNet", class(lasso_mdl))
# measure.bh(lasso_mdl)
# assess.glmnet(lasso_mdl$)

# measure.bh(lasso_mdl, new.x = test_data %>% select(starts_with("mz")) %>% as.matrix,
#            new.y = test_data %>% pull(death3yr))

# lasso_mdl <- glmnet(x = train_data %>% select(starts_with("mz")) %>% as.matrix,
#                     y = train_data %>% pull(death3yr), alpha = 1,
#                     family = "binomial",lambda = lasso_mdl_cv$lambda.min)
# 
# assess.glmnet(lasso_mdl, newx = train_data %>% select(starts_with("mz")) %>% as.matrix, newy = train_data %>% pull(death3yr))
# 
# assess.glmnet(lasso_mdl, newx = test_data %>% select(starts_with("mz")) %>% as.matrix, newy = test_data %>% pull(death3yr))

```

```{r}
library(glmnet)

lasso_mdl_cv <- cv.glmnet(x = train_dat %>% select(gam_scn_res %>% pull(var)) %>% data.matrix,
                    y = train_dat %>% pull(death3yr), alpha = 1,
                    family = "binomial", type.measure = "auc")

lasso_mdl <- glmNet(x = train_dat %>% select(gam_scn_res %>% pull(var)) %>% data.matrix,
                    y = train_dat %>% pull(death3yr), alpha = 1,
                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)

lasso_cv_res <- cv.bh(lasso_mdl, ncv = 10)$measures

# class(lasso_mdl) <- c("glmNet", class(lasso_mdl))
# measure.bh(lasso_mdl)
# assess.glmnet(lasso_mdl$)

# measure.bh(lasso_mdl, new.x = test_data %>% select(gam_scn_res %>% pull(var)) %>% as.matrix,
#            new.y = test_data %>% pull(death3yr))

# lasso_mdl <- glmnet(x = train_data %>% select(starts_with("mz")) %>% as.matrix,
#                     y = train_data %>% pull(death3yr), alpha = 1,
#                     family = "binomial",lambda = lasso_mdl_cv$lambda.min)
# 
# assess.glmnet(lasso_mdl, newx = train_data %>% select(starts_with("mz")) %>% as.matrix, newy = train_data %>% pull(death3yr))
# 
# assess.glmnet(lasso_mdl, newx = test_data %>% select(starts_with("mz")) %>% as.matrix, newy = test_data %>% pull(death3yr))

```

```{r}
library(BHAM)

mz_names <- train_dat %>% select(gam_scn_res %>% pull(var)) %>% names  
sm_df <- data.frame(
  Var = mz_names,
 Func = "s",
 Args ="bs='cr', k=10"
)

sm_obj <- construct_smooth_data(sm_df, train_dat)
```
```{r}
dsn_mat <- sm_obj$data
```

```{r}
# The validation data doesn't overlap with training data
# test_sm_dat <- make_predict_dat(sm_obj$Smooth, dat = valid_dat)

# mdl_defaul <- bmlasso(dsn_mat, train_data$death3yr, family = "binomial", ss  = c(, 0.5))

bmlasso_mdl <- bmlasso(dsn_mat, train_dat$death3yr, family = "binomial",
                           group = make_group(names(dsn_mat)))
```

```{r}
spline_mdl <- bamlasso(dsn_mat, train_dat$death3yr, family = "binomial",
                           group = make_group(names(dsn_mat)))

# bmlasso_cv_res <- cv.bh(bmlasso_mdl, ncv = 10)$measures
# spline_cv_res <- cv.bh(spline_mdl, ncv = 10)$measures

# Training Data Performance
# cat("bmlasso model\n")
# measure.bh(mdl_defaul)
# cat("bmlasso_spline model\n")
# measure.bh(mdl_dflt)

```

```{r eval = F}
test_sm_dat <- BHAM::make_predict_dat(sm_obj$Smooth, dat=test_data)

# Testing Data Performance
cat("bmlasso model\n")
measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=test_data$death3yr)
cat("bmlasso_spline model\n")
measure.bh(mdl_dflt, new.x = test_sm_dat, new.y=test_data$death3yr)
```

# Extremely Penalized Model
```{r}

mdl_pen_extreme <- bamlasso(dsn_mat, train_dat$death3yr, family = "binomial",
                           group = make_group(names(dsn_mat)),
                           ss = c(0.001, 0.5))

# spline_ext_cv_res <- cv.bh(mdl_pen_extreme, ncv = 10)

# Training Data Performance
# measure.bh(mdl_pen_extreme)

# Testing Data Performance
# measure.bh(mdl_pen_extreme, new.x = test_sm_dat, new.y=test_data$death3yr)
```

## Validation

The validation data set have mostly distinct feature from the training data set. Hence, the validation in the traditional sense does not exists.

```{r eval = F}

test_mz_names <- valid_dat  %>% select(starts_with("mz")) %>% names
measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=valid_dat$death3yr)

@@
